---
output: github_document
title:  Example 1 - Generating synthetic data from COVID-19 Cases
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

```

The first step (after installing the package) is to load the library, along with the library dplyr, tidyr and ggplot2.

```{r,echo=T,message=FALSE}
library(dplyr)
library(ggplot2)
library(tidyr)
library(p2synthr)
```

Check the dataset `owid`

```{r,echo=T}
owid
```


Extract cases (from Ireland) as a vector

```{r,echo=T}
irl <- owid |>
        filter(location=="Ireland") |>
        pull(new_cases)

head(irl)
summary(irl)
```
       
Call the function `synth1()` with the following arguments

* The input vector from which the synthetic data will be created
* The argument `group_names` which contains the sub-group names
* The argument `group_prob` which holds the probability of each group bring selected

```{r,echo=T} 
irl_syn_age <- synth1(irl,
                  group_names=c("00-19","20-39","40+"),
                  group_prob = c(0.20,0.40,0.40))

irl_syn_age
sample_n(irl_syn_age,10)
```

For convenience, convert the output to tidy data.
```{r,echo=T}
irl_tidy <- irl_syn_age |>
               pivot_longer(names_to="Age",
                            values_to="Incidence",
                            `00-19`:`40+`)

irl_tidy
```

Plot the results.
```{r,echo=T}
ggplot(irl_tidy,aes(x=Index,y=Incidence,fill=Age))+geom_area()+ggtitle("Synthetic Data")
```
