---
output: github_document
title:  Example 3 - Generating synthetic variant case data
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

```

The first step (after installing the package) is to load the library, along with the library dplyr, ggplot2 and tidyr.

```{r,echo=T,message=FALSE}
library(p2synthr)
library(dplyr)
library(ggplot2)
library(tidyr)
```

Our goal is to create a realistic time series of cases by variant. First, we gather sample case data from our world in data.

```{r,echo=T}
irl <- owid |>
  filter(location=="Ireland") |>
  select(date,new_cases)
irl
```

Next, we call the function `synth_var()`, which breaks out  the cases by variant. The variant dynamics are generated by a cascade of five logistic growth models. The parameters to `synth_variants()` include:

- `cases_ts` - the time series of cases (required)
- `the_dates` - the dates (R date object) for each time series (required)
- `percent_genomic_surveillance_min`, a number between [0,1] indicating the lower percentage of cases sequenced (has defaults)
- `percent_genomic_surveillance_max`, a number between [0,1] indicating the lower percentage of cases sequenced (has defaults)
- `variant_start_time_fractions` a vector of four values indicating at what percentage of overall time a new variant seeds (has defaults)
- `variant_growth_rates`, the growth rates of variants 2 to 5. (has defaults)
- `variant_names`, the variant names (has defaults)
- `setSeed`, dafault = FALSE
- `seedValue`, default = 100


```{r,warning=FALSE,echo=T}
synth_var <- synth_variants(pull(irl,new_cases),
                            pull(irl,date),
                            percent_genomic_surveillance_min = 0.7,
                            percent_genomic_surveillance_max = 0.9)
synth_var
```

The output tibble contains a breakldown of cases by variant (include those that have not been sequenced)

```{r}
glimpse(synth_var)
```

We can now plot the results, by tidying the data.The area plot shows how the synthetic model shows the different dominance stages of the five different viruses.

```{r}
t <- synth_var %>%
       select(Date,TotalCases,starts_with("SynCases"))

t_p <- t %>%
        pivot_longer(names_to="Variant",
                     values_to="Cases",
                     -c(Date,TotalCases))

p1 <- ggplot(t_p,aes(x=Date,y=Cases,fill=Variant))+geom_area()+
      theme(legend.position="top")
p1
```

Finally, the "market share" of the viruses can be shown, and these are driven by the parameters `variant_start_time_fractions` and `variant_growth_rates`.

```{r}
v <- synth_var %>%
  select(Date,ends_with("Fr"))

v_p <- v %>%
  pivot_longer(names_to="Variant",
               values_to="Fraction",
               -c(Date))

p2 <- ggplot(v_p,aes(x=Date,y=Fraction,fill=Variant))+geom_area()+
      theme(legend.position="top")
  
p2

```